We are not the 99 percent: quantifying asphericity in the distribution of Local Group satellites

### 2018-05-08 paper/paper.tex
edit

\begin{figure*}
\centering
\includegraphics[width=0.30\textwidth]{normalized_width_n_dependence.pdf}
\includegraphics[width=0.30\textwidth]{normalized_ca_ratio_n_dependence.pdf}
\includegraphics[width=0.30\textwidth]{normalized_ba_ratio_n_dependence.pdf}
\caption{Normalized asphericity scalars from observations as a function
  of satellite number. Left: plane width; center: $c/a$ ratio;
  right: $b/a$ ratio. 
  The plane width and $c/a$ ratio for the MW is
  consistently found beyond two standard
  deviations.
  M31 shows the opposite trend and is always found
  \emph{within} two standard deviations. \label{fig:normalized_n}} 
\end{figure*}

@@

\subsection{Characterizing Asphericity}

We compare each satellite distribution against its own spherically
randomized distribution.
We keep fixed the radial position of every satellite
with respect to the central galaxy and then randomize its angular
position. 
We repeat this procedure $10^4$ times for each satellite distribution
and proceed to measure the quantities mentioned in the previous section:
$w$, $c/a$ and $b/a$.
This allows us to build a normalized version of all quantities of
interest by subtracting the mean and dividing by the standard
deviation of the values from randomized samples.
We use the normalized quantities to build the analytic probability
distributions for the scalars describing asphericity.

@@

After building the normalized variables with the simulated data we
perform a Kolmogorov-Smirnov test with the null hypothesis of belonging
to a normal distribution with mean and standard deviation computed
from the mean and standard deviation estimated from the data.
Although the physical quantities of interest are bound, unlike the
random variables from a normal distribution, we find that the
distributions for the normalized $w$, $c/a$ and $b/a$ are indeed
consistent with gaussian distributions.

@@

Finally, we use the generated samples to estimate how
common are the deviations from sphericity that we measure in the
observational data.  
We use a double-tailed test around the mean in this comparison, meaning that we
always measure the fraction of points with absolute values larger than the
absolute observed value with the mean substracted.

@@

\begin{figure*}
\centering
\includegraphics[width=0.48\textwidth]{input_illustris1dark_obs_M31_n_11.pdf}
\includegraphics[width=0.48\textwidth]{input_illustris1dark_obs_MW_n_11.pdf}
\caption{Scalars describing satellite shape around the LG galaxies for
  a fixed number of satellites $N_s$=11:
  plane width $w$ in physical units, $c/a$ ratio and $b/a$ ratio.
  Left: M31, right: MW. The square at the center are observations; the
2D histograms are the contour levels after randomizing $10^4$ the
satellite positions while keeping the radial distribution fixed; the
blue circles are the results from Illustris-1-Dark. For the M31 these
three results (observations, randomized observations and simulations)
seem to be in broad agreement, while for the MW the three results
clearly differ. 
The corresponding plots for Illustris-1 and ELVIS are
in Figures \ref{fig:all_plots_illustris1} and
\ref{fig:all_plots_elvis},
respectively. \label{fig:physical_illustris1dark}}     
\end{figure*}

@@

Figure \ref{fig:physical_illustris1dark} summarizes the results in
physical units.
It compares three different groups: observations, simulations and the
observations randomized.
The left panel shows the results for M31 and the panel on the left for
the MW.

@@

This atypicality should be seen as an opportunity to constrain in
great detail the environment that allowed such a pattern to emerge. 
Although broad correlations between LG assembly, pair kinematics, halo
shapes and satellite distributions are expected in LCDM
\citep{2011MNRAS.417.1434F,2014MNRAS.443.1090F,2015ApJ...799...45F,2015MNRAS.452.1052L},
we claim that detailed studies of the satellite asphericity as a
function of halo mass and cosmic web environment are still missing to
understand what features in the initial conditions of our LG are
responsible for the extreme features observed in its satellite
distribution.

@@

\begin{figure*}
\centering
\includegraphics[width=0.37\textwidth]{input_illustris1_obs_M31_n_11.pdf}
\includegraphics[width=0.37\textwidth]{input_illustris1_obs_MW_n_11.pdf}
\includegraphics[width=0.37\textwidth]{input_illustris1_obs_M31_n_11_normed.pdf}
\includegraphics[width=0.37\textwidth]{input_illustris1_obs_MW_n_11_normed.pdf}
\includegraphics[width=0.37\textwidth]{gaussian_model_illustris1_M31_n_11.pdf}
\includegraphics[width=0.37\textwidth]{gaussian_model_illustris1_MW_n_11.pdf}
\caption{Illustris-1 results for the quantities presented for the Illustris1-Dark
  simulation in Figures  
\ref{fig:physical_illustris1dark}, \ref{fig:normalized_illustris1dark}
and \ref{fig:gaussian_illustris1dark}.
Upper row corresponds to the raw values from observations and
simulated pairs, while the second row normalizes the same values to
the mean and standard deviation on its spherically randomized
counterparts. 
\label{fig:all_plots_illustris1}}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=0.37\textwidth]{input_elvis_obs_M31_n_11.pdf}
\includegraphics[width=0.37\textwidth]{input_elvis_obs_MW_n_11.pdf}
\includegraphics[width=0.37\textwidth]{input_elvis_obs_M31_n_11_normed.pdf}
\includegraphics[width=0.37\textwidth]{input_elvis_obs_MW_n_11_normed.pdf}
\includegraphics[width=0.37\textwidth]{gaussian_model_elvis_M31_n_11.pdf}
\includegraphics[width=0.37\textwidth]{gaussian_model_elvis_MW_n_11.pdf}
\caption{ELVIs results for the quantities presented for the Illustris1-Dark
  simulation in Figures  \ref{fig:scatter_width},
  \ref{fig:scatter_ca_ratio}, \ref{fig:scatter_ba_ratio}.
Upper row corresponds to the raw values from observations and
simulated pairs, while the second row normalizes the same values to
the mean and standard deviation on its spherically randomized
counterparts. 
\label{fig:all_plots_elvis}}
\end{figure*}

>>>>

\begin{figure*}
\centering
\includegraphics[width=0.30\textwidth]{normalized_width_n_dependence.pdf}
\includegraphics[width=0.30\textwidth]{normalized_ca_ratio_n_dependence.pdf}
\includegraphics[width=0.30\textwidth]{normalized_ba_ratio_n_dependence.pdf}
\caption{Asphericity scalars from observations as a function
  of satellite number. 
  Left: plane width; center: $c/a$ ratio;
  right: $b/a$ ratio. 
  These quantities can take positive or negative values because they
  were recentered to the mean and normalized to the standard deviation 
  of the spherically randomized satellites as explained in Section
  \ref{sec:normalization}. 
  The plane width and $c/a$ ratio for the MW is
  consistently found \emph{beyond} two standard
  deviations of the expected result for an statistically spherical
  distribution. 
  M31 shows the opposite trend and is always found
  \emph{within} two standard deviations. \label{fig:normalized_n}} 
\end{figure*}

@@

\subsection{Characterizing asphericity with normalized scalars}
\label{sec:normalization}
We compare each satellite distribution against its own spherically
randomized distribution.
We keep fixed the radial position of every satellite
with respect to the central galaxy and then randomize its angular
position. 
We repeat this procedure $10^4$ times for each satellite distribution
and proceed to measure the quantities mentioned in the previous section:
$w$, $c/a$ and $b/a$.
This allows us to build a normalized version of all quantities of
interest by subtracting the mean and dividing by the standard
deviation of the values from randomized samples.
We use the normalized quantities to build the analytic probability
distributions for the scalars describing asphericity.

@@

After building the normalized variables for the simulated data we perform a
Kolmogorov-Smirnov test with the null hypothesis of belonging to a
normal distribution with mean and standard deviation computed from the
mean and standard deviation estimated from the data. 
Although the physical quantities of interest are bound, we find that the
distributions for the normalized $w$, $c/a$ and $b/a$ are indeed
consistent with gaussian distributions.

@@

Finally, we use the generated samples to estimate how
common are the deviations from sphericity that we measure in the
observational data.  
We use a double-tailed test around the mean in this comparison,
meaning that we count the fraction of generated points points with
absolute values larger than the absolute observed value with the mean
substracted.

@@

\begin{figure*}
\centering
\includegraphics[width=0.48\textwidth]{input_illustris1dark_obs_M31_n_11.pdf}
\includegraphics[width=0.48\textwidth]{input_illustris1dark_obs_MW_n_11.pdf}
\caption{Scalars in physical units describing satellite shape around the LG galaxies for
  a fixed number of satellites $N_s$=11: 
  plane width $w$, $c/a$ ratio and $b/a$ ratio.
  Left: M31, right: MW. 
  The histograms along the diagonal show the distribution for the
  $10^4$ values after spherical randomization; the vertical dashed
  lines show the 16, 50 and 84 percentiles of this distribution; the vertical solid line correspond to
  observations. 
  Outside the diagonal there are 2D histograms showing the $1$, $2$, $3$
  and $4\sigma$ contours for the randomized data.  
  The square at the center of the cross are observations.
  The blue circles are the results from Illustris-1-Dark. 
  For the M31 these three results
  (observations, randomized observations and simulations) seem to be
  in broad agreement, while for the MW they clearly differ. 
  The corresponding plots for Illustris-1 and ELVIS are
  in Figures \ref{fig:all_plots_illustris1} and
\ref{fig:all_plots_elvis}, respectively. \label{fig:physical_illustris1dark}}
\end{figure*}

@@

Figure \ref{fig:physical_illustris1dark} summarizes the results in
physical units.
It compares three different groups: observations, simulations and the
observations randomized.
The left panel shows the results for M31 and the panel on the right for
the MW.

@@

The LG atypicality should be seen as an opportunity to constrain in
great detail the environment that allowed such a pattern to emerge. 
Although broad correlations between LG assembly, pair kinematics, halo
shapes and satellite distributions are expected in LCDM
\citep{2011MNRAS.417.1434F,2014MNRAS.443.1090F,2015ApJ...799...45F,2015MNRAS.452.1052L},
we claim that detailed studies of the satellite asphericity as a
function of halo mass and cosmic web environment are still missing to
understand what features in the initial conditions of our LG are
responsible for the extreme features observed in its satellite
distribution.

@@

\begin{figure*}
\centering
\includegraphics[width=0.37\textwidth]{input_illustris1_obs_M31_n_11.pdf}
\includegraphics[width=0.37\textwidth]{input_illustris1_obs_MW_n_11.pdf}
\includegraphics[width=0.37\textwidth]{input_illustris1_obs_M31_n_11_normed.pdf}
\includegraphics[width=0.37\textwidth]{input_illustris1_obs_MW_n_11_normed.pdf}
\includegraphics[width=0.37\textwidth]{gaussian_model_illustris1_M31_n_11.pdf}
\includegraphics[width=0.37\textwidth]{gaussian_model_illustris1_MW_n_11.pdf}
\caption{Illustris-1 results for the same quantities presented for the Illustris-1-Dark
  simulation in Figures  
\ref{fig:physical_illustris1dark}, \ref{fig:normalized_illustris1dark}
and \ref{fig:gaussian_illustris1dark}.
Upper row corresponds to the raw values from observations and
simulated pairs, while the second row normalizes the same values to
the mean and standard deviation on its spherically randomized
counterparts. 
\label{fig:all_plots_illustris1}}
\end{figure*}

\begin{figure*}
\centering
\includegraphics[width=0.37\textwidth]{input_elvis_obs_M31_n_11.pdf}
\includegraphics[width=0.37\textwidth]{input_elvis_obs_MW_n_11.pdf}
\includegraphics[width=0.37\textwidth]{input_elvis_obs_M31_n_11_normed.pdf}
\includegraphics[width=0.37\textwidth]{input_elvis_obs_MW_n_11_normed.pdf}
\includegraphics[width=0.37\textwidth]{gaussian_model_elvis_M31_n_11.pdf}
\includegraphics[width=0.37\textwidth]{gaussian_model_elvis_MW_n_11.pdf}
\caption{ELVIS results for the same quantities presented for the Illustris1-Dark
  simulation in Figures  
\ref{fig:physical_illustris1dark}, \ref{fig:normalized_illustris1dark}
and \ref{fig:gaussian_illustris1dark}.
Upper row corresponds to the raw values from observations and
simulated pairs, while the second row normalizes the same values to
the mean and standard deviation on its spherically randomized
counterparts. 
\label{fig:all_plots_elvis}}
\end{figure*}
